Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-06-08 Master Thesis/synopsis.tex
Complete argumentation in remark 3.3
Prefer ' notation instead over sigma notation

\begin{proof}
We adopt the notation from the previous proof. First, we shall prove that the following map is injective for every model $\mathfrak{X}$ of $K$:
\[ f: \mathrm{H}^3(K) \longrightarrow \prod_{y\in\mathfrak{X}_0}\mathrm{H}^3(K_y) \]
For $x\in\mathfrak{X}$ let $\mathcal{O}_x$ be a henselization of $\mathcal{O}_{\mathfrak{X}, x}$ and $K_x$ its quotient field. For $y\in\mathfrak{X}_0$ set $P_y=\operatorname{Spec}(\mathcal{O}_y)_1$. For $z\in P_y$ let $\mathcal{O}_z$ be a henselization of $\mathcal{O}_y$ at $z$ and $K_z$ its quotient field. For $y\in\mathfrak{X}_0$ and $x\in\mathfrak{X}_1$ let $P_{x,y}$ be the subset of $P_y$ consisting of all $z\in P_y$ which are mapped to $x$ under $\operatorname{Spec}(\mathcal{O}_y)\to\mathfrak{X}$. Consider the following commutative diagram:
\[ \begin{tikzcd}
\mathrm{H}^3(K)\ar[r, "\varphi"]\ar[d, "f"'] & \displaystyle\prod_{x\in\mathfrak{X}_1}\mathrm{H}^3(K_x)\ar[d, "\psi"]\\
\displaystyle\prod_{y\in\mathfrak{X}_0}\mathrm{H}^3(K_y) \ar[r, "g"']& \displaystyle\prod_{\substack{x\in\mathfrak{X}_1 \\ y\in\mathfrak{X}_0}}\prod_{z\in P_{x,y}}\mathrm{H}^3(K_z)
\end{tikzcd} \]
According to [Sa], Theorem (2-15) $g$ is injective. Now, $\psi\circ\varphi$ can be identified with the injective map $\Phi$ from the proof of the previous Lemma. Thus its image is contained in $\bigoplus\mathrm{H}^3(K_w)$. Therefore, $f$ is injective as well and its image is contained in $\bigoplus\mathrm{H}^3(K_y)$. Now, one easily checks that:
\[ \operatorname{supp}(f) = \bigcap_{\mathfrak{X}\in \mathfrak{M}(K)} \pi_\mathfrak{X}^{-1}(\{ y\in\mathfrak{X}_0\mid \operatorname{res}_{K_y/K}(f)\neq 0 \}) \]
Note that the right set is closed in $\operatorname{Spv}(K)$.
\end{proof}

@@

\begin{remark}\label{3.3}
If $L\subset M$ are intermediate fields of $\overline{K}/K$, which are finite over $K$, then $\sigma$ also induces bijections
\[ \operatorname{Spv}_\text{d}^{(1)}(L) \to \operatorname{Spv}_\text{d}^{(1)}(L'),\qquad \operatorname{Spv}_\text{d}^{(1)}(M) \to \operatorname{Spv}_\text{d}^{(1)}(M') \]
such that the following diagram commutes:
\[\begin{tikzcd}
\operatorname{Spv}_\text{d}^{(1)}(M) \rar["\sigma"]\dar & \operatorname{Spv}_\text{d}^{(1)}(M')\dar\\
\operatorname{Spv}_\text{d}^{(1)}(L) \rar["\sigma"'] & \operatorname{Spv}_\text{d}^{(1)}(L')
\end{tikzcd}\]
where the vertical maps are restrictions. Let $w\in\operatorname{Spv}_\text{d}^{(1)}(M),\ v=w|_L$ and $w'=\sigma(w),\ v'=\sigma(v)$. Then the following diagram commutes:
\[ \begin{tikzcd}
k(w') \rar["\varphi_w"] & k(w)\\
k(v') \rar["\varphi_v"']\uar[hook] & k(v)\uar[hook]
\end{tikzcd} \]
By limit process we see that $\sigma$ induces a bijection $\sigma: \operatorname{Spv}_\text{d}^{(1)}(\overline{K})\to \operatorname{Spv}_\text{d}^{(1)}(\overline{K'})$. Let $\overline{v}\in \operatorname{Spv}_\text{d}^{(1)}(\overline{K})$ and $G_{\overline{v}}\subset G_K$ be the decomposition group of $\overline{v}$. Then we have $\sigma(G_{\overline{v}}) = G_{\sigma(\overline{v})}$. Now it is easy to see that for every $g\in G_K$:
\[ \sigma(g\overline{v}) = \sigma(g)\sigma(\overline{v}) \]
\end{remark}

@@

Before we start proving Pop's Theorem \ref{thm:main-result}, we need some tools that will help us constructing a group isomorphism $K'^\times\to K^\times$ from the modulo $n$ isomorphisms $K'^\times/n\to K^\times/n$.

@@

For the existence of $\varphi$, let $(S,T)$ be an admissible pair. First, we show that there is a unique homomorphism $\varphi^{S,T}$, making the following diagram commutative:
\[ (\star) \qquad \begin{tikzcd}
K'^{S'} \rar["\varphi^{S,T}", dashed] \dar["r_{S',T'}"'] & K^S \dar["r_{S,T}"]\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times \rar["\prod\varphi_P"'] & \displaystyle \prod_{P\in T}k(P)^\times
\end{tikzcd} \phantom{\qquad (\star)} \]
The vertical maps are injective by definition. Thus, it suffices to show that the following composition is trivial:
\[\varrho:K'^{S'}\longrightarrow\prod_{P\in T} k(\sigma P)^\times \longrightarrow \prod_{P\in T} k(P)^\times\longrightarrow\operatorname{coker}(r_{S,T})\] 
Consider the commutative diagram which results from tensoring $(\star)$ with $\mathbb{Z}/n\mathbb{Z}$:
\[ \begin{tikzcd}
K'^{S'}/n \rar[dashed]\dar & K^S/n\dar\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times/n \rar & \displaystyle \prod_{P\in T}k(P)^\times/n
\end{tikzcd} \]
According to Lemma~\ref{4.3} and Lemma~\ref{4.5} (ii) this diagram can now be completed by $\varphi_n: K'^{S'}/n\to K^S/n$. This implies $\varrho/n=0$, i.e. $\operatorname{im}(\varrho)\subset n\operatorname{coker}(r_{S,T})$ for every $n\in\mathbb{N}$. By Lemma~\ref{5.2} (iii) and Lemma~\ref{5.4} it follows $\varrho = 0$. By symmetry, $\varphi^{S,T}$ is an isomorphism.

>>>>

\begin{proof}
We adopt the notation from the previous proof. First, we shall prove that the following map is injective for every model $\mathfrak{X}$ of $K$:
\[ f: \mathrm{H}^3(K) \longrightarrow \prod_{y\in\mathfrak{X}_0}\mathrm{H}^3(K_y) \]
For $x\in\mathfrak{X}$ let $\mathcal{O}_x$ be a henselization of $\mathcal{O}_{\mathfrak{X}, x}$ and $K_x$ its quotient field. For $y\in\mathfrak{X}_0$ set $P_y=\operatorname{Spec}(\mathcal{O}_y)_1$. For $z\in P_y$ let $\mathcal{O}_z$ be a henselization of $\mathcal{O}_y$ at $z$ and $K_z$ its quotient field. For $y\in\mathfrak{X}_0$ and $x\in\mathfrak{X}_1$ let $P_{x,y}$ be the subset of $P_y$ consisting of all $z\in P_y$ which are mapped to $x$ under $\operatorname{Spec}(\mathcal{O}_y)\to\mathfrak{X}$. Consider the following commutative diagram:
\[ \begin{tikzcd}
\mathrm{H}^3(K)\ar[r, "\varphi"]\ar[d, "f"'] & \displaystyle\prod_{x\in\mathfrak{X}_1}\mathrm{H}^3(K_x)\ar[d, "\psi"]\\
\displaystyle\prod_{y\in\mathfrak{X}_0}\mathrm{H}^3(K_y) \ar[r, "g"']& \displaystyle\prod_{\substack{x\in\mathfrak{X}_1 \\ y\in\mathfrak{X}_0}}\prod_{z\in P_{x,y}}\mathrm{H}^3(K_z)
\end{tikzcd} \]
According to [Sa], Theorem (2-15) $g$ is injective. Now, $\psi\circ\varphi$ can be identified with the injective map $\Phi$ from the proof of the previous Lemma. Thus its image is contained in $\bigoplus\mathrm{H}^3(K_w)$. Therefore, $f$ is injective as well and its image is contained in $\bigoplus\mathrm{H}^3(K_y)$. Now, one checks that:
\[ \operatorname{supp}(f) = \bigcap_{\mathfrak{X}\in \mathfrak{M}(K)} \pi_\mathfrak{X}^{-1}(\{ y\in\mathfrak{X}_0\mid \operatorname{res}_{K_y/K}(f)\neq 0 \}) \]
Note that the right set is closed in $\operatorname{Spv}(K)$.
\end{proof}

@@

\begin{remark}\label{3.3}
If $L\subset M$ are intermediate fields of $\overline{K}/K$, which are finite over $K$, then $\sigma$ also induces bijections
\[ \operatorname{Spv}_\text{d}^{(1)}(L) \to \operatorname{Spv}_\text{d}^{(1)}(L'),\qquad \operatorname{Spv}_\text{d}^{(1)}(M) \to \operatorname{Spv}_\text{d}^{(1)}(M') \]
such that the following diagram commutes:
\[\begin{tikzcd}
\operatorname{Spv}_\text{d}^{(1)}(M) \rar["\sigma"]\dar & \operatorname{Spv}_\text{d}^{(1)}(M')\dar\\
\operatorname{Spv}_\text{d}^{(1)}(L) \rar["\sigma"'] & \operatorname{Spv}_\text{d}^{(1)}(L')
\end{tikzcd}\]
where the vertical maps are restrictions. Let $w\in\operatorname{Spv}_\text{d}^{(1)}(M),\ v=w|_L$ and $w'=\sigma(w),\ v'=\sigma(v)$. Then the following diagram commutes:
\[ \begin{tikzcd}
k(w') \rar["\varphi_w"] & k(w)\\
k(v') \rar["\varphi_v"']\uar[hook] & k(v)\uar[hook]
\end{tikzcd} \]
By limit process we see that $\sigma$ induces a bijection $\sigma: \operatorname{Spv}_\text{d}^{(1)}(\overline{K})\to \operatorname{Spv}_\text{d}^{(1)}(\overline{K'})$. Let $\overline{v}\in \operatorname{Spv}_\text{d}^{(1)}(\overline{K})$ and $G_{\overline{v}}\subset G_K$ be the decomposition group of $\overline{v}$. Then we have $\sigma(G_{\overline{v}}) = G_{\sigma(\overline{v})}$. It follows for any $g\in G_K$:
\[G_{\sigma(g\overline{v})} = \sigma(G_{g\overline{v}}) = \sigma(g^{-1} G_{\overline{v}}g) = \sigma(g)^{-1} G_{\sigma(\overline{v})}\sigma(g) = G_{\sigma(g)\sigma(\overline{v})}\]
Therefore, $\sigma(g\overline{v})=\sigma(g)\sigma(\overline{v})$.
\end{remark}

@@

Before we start proving Pop's Theorem \ref{thm:main-result}, we need some tools that will help us constructing a group isomorphism $K'^\times\to K^\times$ from the modulo $n$ isomorphisms $\varphi_n: K'^\times/n\to K^\times/n$.

@@

For the existence of $\varphi$, let $(S,T)$ be an admissible pair. First, we show that there is a unique homomorphism $\varphi^{S,T}$, making the following diagram commutative:
\[ (\star) \qquad \begin{tikzcd}
K'^{S'} \rar["\varphi^{S,T}", dashed] \dar["r_{S',T'}"'] & K^S \dar["r_{S,T}"]\\
\displaystyle \prod_{P\in T}k(P')^\times \rar["\prod\varphi_P"'] & \displaystyle \prod_{P\in T}k(P)^\times
\end{tikzcd} \phantom{\qquad (\star)} \]
The vertical maps are injective by definition. Thus, it suffices to show that the following composition is trivial:
\[\varrho:K'^{S'}\longrightarrow\prod_{P\in T} k(P')^\times \longrightarrow \prod_{P\in T} k(P)^\times\longrightarrow\operatorname{coker}(r_{S,T})\] 
Consider the commutative diagram which results from tensoring $(\star)$ with $\mathbb{Z}/n\mathbb{Z}$:
\[ \begin{tikzcd}
K'^{S'}/n \rar[dashed]\dar & K^S/n\dar\\
\displaystyle \prod_{P\in T}k(P')^\times/n \rar & \displaystyle \prod_{P\in T}k(P)^\times/n
\end{tikzcd} \]
According to Lemma~\ref{4.3} and Lemma~\ref{4.5} (ii) this diagram can now be completed by $\varphi_n: K'^{S'}/n\to K^S/n$. This implies $\varrho/n=0$, i.e. $\operatorname{im}(\varrho)\subset n\operatorname{coker}(r_{S,T})$ for every $n\in\mathbb{N}$. By Lemma~\ref{5.2} (iii) and Lemma~\ref{5.4} it follows $\varrho = 0$. By symmetry, $\varphi^{S,T}$ is an isomorphism.
